---
title: "Mapping Geospatial Networks"
---

# LIBRARIES
```{r setup, include=FALSE}
library(osmdata)    # Open Street Map Overpass API
library(osrm)       # Open Street Map Routing API

library(sf)
library(nngeo)
library(mapview)

library(tidyverse)
library(tidyquant)
```




# 0.0 DATA
## Import FL Listings
```{r}
broward_listings_tbl <- read_rds("C:\\Users\\DavidStephens\\Desktop\\Github\\geospatial_analysis\\00_data\\listings_filtered_dt.rds") %>%
  select(id, longitude, latitude)
```




# 1.0 OSM OVERPASS API
```{r}
# - Get roads, water, etc
# - Resource: # https://wiki.openstreetmap.org/wiki/Map_features

broward_bbox <- getbb("Broward County, FL")
?osmdata()


# tags
available_tags("highway") # https://wiki.openstreetmap.org/wiki/Map_features#Roads
```




## Getting Smaller roads
```{r}
# medium streets
broward_medium_streets_sf <- opq(broward_bbox) %>%
    add_osm_feature(
        key   = "highway", 
        value = c("secondary", "tertiary", "secondary_link", "tertiary_link")
    ) %>%
    osmdata_sf()

broward_medium_streets_sf %>% write_rds("00_data/04_broward_medium_streets_sf.rds")
broward_medium_streets_sf <- read_rds("00_data/04_broward_medium_streets_sf.rds")



# visualize
mapview(
    broward_medium_streets_sf$osm_lines, 
    color = "darkgreen",
    layer.name = "Streets"
) +
    mapview(
        broward_highways_sf$osm_lines,
        layer.name = "Highways",
        color = "purple"
    )
```


## Residential
x.x Not Run: too much detail for this analysis
```{r, eval=FALSE}
# residential
broward_residential_streets_sf <- opq(broward_bbox) %>%
    add_osm_feature(
        key   = "highway", 
        value = c("residential")
    ) %>%
    osmdata_sf()


broward_residential_streets_sf %>% write_rds("00_data/04_broward_residential_streets_sf.rds")
broward_residential_streets_sf <- read_rds("00_data/04_broward_residential_streets_sf.rds")
```





# 2.0 CUSTOMERS & WAREHOUSES



Read in customers & distributors
```{r}
customer_sample_set    <- read_rds("00_data/04_networks_broward_customer_sample_set.rds")
distributor_sample_set <- read_rds("00_data/04_networks_broward_distributor_sample_set.rds")
```


## Visualize
```{r}
# * Customer Data ----
broward_customers_sf <- customer_sample_set %>%
    rowid_to_column(var = "customer_id")

# * Warehouse Data ----
broward_distributors_sf <- distributor_sample_set %>%
    rowid_to_column(var = "distributor_id")



# visualize
mapview(
    broward_medium_streets_sf$osm_lines, 
    color      = "darkgreen",
    layer.name = "Streets",
    lwd        = 0.5 
) +
    mapview(
        broward_highways_sf$osm_lines,
        layer.name = "Highways",
        color      = "purple",
        lwd        = 2
    ) +
    mapview(
        broward_customers_sf,
        col.region = "blue",
        color      = "white",
        layer.name = "Customers",
        cex        = 12
    ) +
    mapview(
        broward_distributors_sf,
        col.region = "magenta",
        color      = "white",
        layer.name = "Warehouses",
        cex        = 20
    )
```




# 3.0 NEAREST NEIGHBORS
## Getting Nearest Neighbors with nngeo
Alternatively we can use sfnetworks
```{r}
network_broward_ids <- st_nn(
    x = broward_distributors_sf,
    y = broward_customers_sf,
    k = nrow(broward_customers_sf),
    progress = T
)


network_broward_lines_sf <- st_connect(
    x   = broward_distributors_sf,
    y   = broward_customers_sf,
    ids = network_broward_ids
)



# visualize
mapview(
    broward_medium_streets_sf$osm_lines, 
    color      = "darkgreen",
    layer.name = "Streets",
    lwd        = 0.5 
) +
    mapview(
        broward_highways_sf$osm_lines,
        layer.name = "Highways",
        color      = "purple",
        lwd        = 2
    ) +
    mapview(
        broward_customers_sf,
        col.region = "blue",
        color      = "white",
        layer.name = "Customers",
        cex        = 12
    ) +
    mapview(
        broward_distributors_sf,
        col.region = "magenta",
        color      = "white",
        layer.name = "Warehouses",
        cex        = 20
    ) +
    mapview(
        network_broward_lines_sf,
        color      = "red"
    )
```


## Approximate Shortest Path
```{r}
nodes_broward_tbl <- network_broward_ids %>%
    enframe(
        name  = "distributor_id",
        value = "customer_id"
    ) %>%
    unnest(customer_id)


shortest_broward_network_sf <- network_broward_lines_sf %>%
    st_bind_cols(nodes_broward_tbl) %>%
    mutate(len = st_length(geometry)) %>%
    relocate(len, .after = customer_id) %>%
    
    group_by(customer_id) %>%
    filter(len == min(len)) %>%
    ungroup()



# visualize
mapview(
    broward_medium_streets_sf$osm_lines,
    color      = "darkgreen",
    layer.name = "Streets",
    lwd        = 0.5
) +
    mapview(
        broward_highways_sf$osm_lines,
        layer.name = "Highways",
        color      = "purple",
        lwd        = 2
    ) +
    mapview(
        broward_customers_sf,
        col.region = "blue",
        color      = "white",
        layer.name = "Customers",
        cex        = 12
    ) +
    mapview(
        broward_distributors_sf,
        col.region = "magenta",
        color      = "white",
        layer.name = "Warehouses",
        cex        = 20
    ) +
    mapview(
        shortest_broward_network_sf,
        color      = "red",
        layer.name = "Shortest Network"
    )
```


## Defining the Trip Points
```{r}
broward_route_points_sf <- broward_distributors_sf %>%
    
  bind_rows(broward_customers_sf) %>%
  select(type, distributor_id, customer_id, everything()) %>%
  
  # Adding in the distributor that the customer belongs to
  left_join(
      shortest_broward_network_sf %>% 
          select(distributor_id, customer_id) %>%
          as_tibble() %>%
          rename(distributor_to = distributor_id) %>%
          select(-geometry),
      by   = "customer_id"
  ) %>%
  
  # Cleanup distributor_to
  mutate(distributor_to = ifelse(is.na(distributor_to), distributor_id, distributor_to)) %>%
  mutate(distributor_to = as.factor(distributor_to))



# visualize
mapview(
    broward_route_points_sf,
    zcol       = "distributor_to",
    layer.name = "Distributor Network"
)
```




# 4.0 OSM ROUTES API
## * Getting 1 route
```{r}
route_1_list <- broward_route_points_sf %>%
  filter(distributor_to == 1) %>%
  osrmTrip()

route_1_list[[1]]$summary

mapview(route_1_list[[1]]$trip)
```


## Mapping to many routes
```{r}
warehouse_trips_tbl <- broward_route_points_sf %>%
  group_by(distributor_to) %>%
  group_nest() %>%
  mutate(trip = map(data, .f = osrmTrip))

warehouse_trips_tbl$trip[[1]]

warehouse_trips_sf <- warehouse_trips_tbl %>%
    
  # Double unnest
  select(-data) %>%
  unnest(trip) %>%
  unnest(trip) %>%
  
  # Get first item
  group_by(distributor_to) %>%
  slice(1) %>%
  ungroup() %>%
  
  # Unnest sf object and convert to sf
  unnest(trip) %>%
  st_as_sf()




# visualize routes
mapview(
  broward_customers_sf,
  col.region = "blue",
  color      = "white",
  layer.name = "Customers",
  cex        = 12
) +
  mapview(
      broward_distributors_sf,
      col.region = "magenta",
      color      = "white",
      layer.name = "Warehouses",
      cex        = 20
  ) +
  mapview(
      warehouse_trips_sf,
      zcol = "distributor_to",
      color = tidyquant::palette_dark()[c(1,2,4)],
      layer.name = "Trip"
  )
```




# 5.0 COSTS
## Estimating trip cost
$2.90 cost per mile from National Private Truck Council (NPTC)
```{r}
warehouse_trips_tbl %>%
  
  # Double unnest
  select(-data) %>%
  unnest(trip) %>%
  unnest(trip) %>%
  
  # Get the 2nd item
  group_by(distributor_to) %>%
  slice(2) %>%
  ungroup() %>%
  
  # Trick: Unnest wider
  unnest_wider(trip) %>%
  
  # Add our costs
  mutate(distance = distance * 0.621371,
         driver_cost_per_trip = 500,
         cost_per_mile = 2.90,
         total_cost  = distance * cost_per_mile + driver_cost_per_trip) %>%
  rename(duration_min = duration,
         distance_miles = distance)
```































